### 2016-07-11 PALSE/talk_Kyoto.tex
start Counterion condensation

\documentclass[xcolor=table]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage[super]{nth}
%\usetheme{Boadilla}
%\usecolortheme{rose}
%\usecolortheme{crane}
%\usefonttheme{structuresmallcapsserif}
%\setbeamertemplate{navigation symbols}{}

\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
\definecolor{Accent1}{rgb}{0.76,0.36,0.13}
\definecolor{Accent2}{rgb}{0.54,0.1,0.4}

\mode<presentation>{\usetheme{ilm}}
%\usecolortheme{rose}
%\useinnertheme[shadow]{circles}
%\usecolortheme{whale}
%\useoutertheme{infolines}

%\usecolortheme[named=Accent1]{structure}

%\setbeamerfont{page number in head/foot}{size=\large}
%\setbeamercolor{page number in head/foot}{fg=Main}
%% page/total
%%\setbeamertemplate{footline}[frame number]
%% pas de total
%\setbeamertemplate{footline}{%
%    	\hfill%
%	\usebeamercolor[fg]{page number in head/foot}%
%	\usebeamerfont{page number in head/foot}%
%	\insertframenumber\kern1em\vskip2pt%
%}
%
%\setbeamersize{text margin left=1em}
%\setbeamersize{text margin right=1em}

\usepackage[overlay,absolute]{textpos}
\setlength{\TPHorizModule}{10mm}
\setlength{\TPVertModule}{\TPHorizModule}
\textblockorigin{10mm}{10mm} % start everything near the top-left corner
\setlength{\parindent}{0pt}

%font
\usepackage[T1]{fontenc}
\usepackage{times}
%\usepackage[oldstylenums]{kpfonts}

%\include{macros}
\usepackage{ifthen}

%proper math and math symbols
\usepackage{amsmath,mathtools}
\usepackage{amssymb}

\usepackage{datenumber,fp}

\usepackage[separate-uncertainty = true]{siunitx}

\usepackage{tabu}
\usepackage{multirow}
\usepackage{booktabs}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, decorations.text, shapes.geometric, matrix,patterns,mindmap}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.9}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}
\makeatletter
\newcommand*{\overlaynumber}{\number\beamer@slideinframe}
\tikzset{
  beamer externalizing/.style={%
    execute at end picture={%
      \tikzifexternalizing{%
        \ifbeamer@anotherslide
        \pgfexternalstorecommand{\string\global\string\beamer@anotherslidetrue}%
        \fi
      }{}%
    }%
  },
  external/optimize=false
}
\let\orig@tikzsetnextfilename=\tikzsetnextfilename
\renewcommand\tikzsetnextfilename[1]{\orig@tikzsetnextfilename{#1-\overlaynumber}}
\makeatother

\tikzset{every picture/.style={beamer externalizing}}
\tikzexternalize
\tikzsetexternalprefix{fig_presentation/}
%\tikzset{external/optimize=false}
%\tikzset{external/force remake}

%link or play movies
\usepackage{multimedia}

%chemistry
\usepackage[version=3]{mhchem}
\usepackage{chemfig}
\usepackage{setspace}

%beamer related package

\usepackage{todonotes}
\presetkeys{todonotes}{inline}{}

\tikzset{onslide/.code args={<#1>#2}{%
  \only<#1>{\pgfkeysalso{#2}} %
}}%

%bibliography
\usepackage[style=authoryear-comp, language=british,eprint=false, url=false, doi=false, sortcites=true, sorting=none, isbn=false, firstinits=true,maxcitenames=6]{biblatex}
%minimal citations
\AtEveryCitekey{%
	\clearfield{title}
	\clearfield{pages}
	\clearfield{volume}
	\clearfield{number}
	\clearfield{month}}
\newcommand{\myfullcite}[1]{{\scriptsize\fullcite{#1}}}
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}
%\bibliography{biblio}

\newcolumntype{P}[1]{>{\raggedright}p{#1}}

\institute[iLM]{Univ Lyon, Université Claude Bernard Lyon 1, CNRS, Institut Lumière Matière}
\title[processionary gel]{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}
\author[M. Leocmach]{Mathieu Leocmach {\usebeamerfont{normal text}\texttt{\usebeamercolor[fg]{normal text}\footnotesize @LamSonLeoc}}\vspace{-\baselineskip}}
\date{%10 August 2016
}
\titlegraphic{\vspace{-\baselineskip}
	\begin{tabu}{X[c]X[c]X[c]X[c]}
		\multicolumn{2}{c}{\usebeamerfont{institute}\usebeamercolor[fg]{institute}ENS de Lyon, Laboratoire de Chimie}&
		\multicolumn{2}{c}{\usebeamerfont{institute}\usebeamercolor[fg]{institute}ENS de Lyon, Laboratoire de Physique}\\\cmidrule(r){1-2}\cmidrule(l){3-4}
		\includegraphics[height=3\baselineskip]{presentation/Hassan}&
		\includegraphics[height=3\baselineskip]{presentation/Cyrille}&
		\includegraphics[height=3\baselineskip]{presentation/Nicolas_Taberlet}&
		\includegraphics[height=3\baselineskip]{../Yaourt/Seb}\\
		Hassan & Cyrille & Nicolas & S\'{e}bastien\\
		Srour & Monnereau & Taberlet & Manneville\\\cmidrule(r){1-2}
		\multicolumn{2}{c}{Martien Duvall Deffo Ayagou}&\\
		\multicolumn{2}{c}{Thi Thanh-Tam Nguyen}&\raisebox{-0.1em}[0pt][0pt]{\includegraphics[height=2\baselineskip]{presentation/invest-avenir.jpg}}&\raisebox{-0.1em}[0pt][0pt]{\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{../Yaourt/NEW-Logo-ERC-OUTLINE}}\\
	\end{tabu}

\bigskip
Martien Duvall Deffo Ayagou, Thi Thanh-Tam Nguyen%, Chantal Andraud

	%\vfill
	%\includegraphics[height=2\baselineskip]{logo_ens-lyon}\quad
	%\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
	%\includegraphics[height=2\baselineskip]{../../Yaourt/CNRSfilaire-Q}\quad
	%\includegraphics[height=2\baselineskip]{CRPP}\quad
	%\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}
	}

\newlength{\mylength}

%\includeonly{creep_beamer}

\begin{document}
\tikzset{every mark/.append style={scale=0.8}}
\pgfplotsset{every axis/.append style={footnotesize}}

\pgfplotscreateplotcyclelist{earthy}{%
{red!40!black,mark=o},
{red!60!black,mark=triangle, every mark/.append style={rotate=180}},
{red!80!black,mark=square},
{red,mark=triangle},
{red!80!yellow, mark=diamond},
red!60!yellow,
red!40!yellow,
}

\AtBeginSection[]{
	\addtocounter{framenumber}{-1}
	\begin{frame}[plain]
		\tableofcontents[currentsection, hideothersubsections]
	\end{frame}
}

\begin{frame}{\pgfuseimage{cnrs-logo}\hspace*{0.3cm}\pgfuseimage{ucbl-logo}\pgfuseimage{univlyon-logo}}%[plain]
	\titlepage
\end{frame}

\setcounter{framenumber}{0}

\begin{frame}{Fine chemistry}
	\vspace{\baselineskip}
	\tikzsetnextfilename{polymerisation}%
	\input{fig_presentation/polymerisation.pgf}

	\bigskip
	\tikzsetnextfilename{polymerisation_PImBr}%
	\input{fig_presentation/polymerisation_PImBr.pgf}
\end{frame}

\begin{frame}{Physical polyelectrolytes hydrogel}
	\tikzsetnextfilename{polyelectrolytes}%
\input{fig_presentation/polyelectrolytes.pgf}
\begin{columns}
\column{0.5\textwidth}
\structure{Reversible gel}
\begin{itemize}
	\item in deionised, neutral water
	\item yield stress fuild
\end{itemize}
\structure{Solution} if
\begin{itemize}
	\item ionic strength, pH
	\item non ionic head
\end{itemize}
$\Rightarrow$ Head-to-body ionic bonds

\column{0.5\textwidth}
\hfill Herschel-Buckley law\\
\hfill $\sigma =\sigma_c + A \dot{\gamma}^n$
\tikzsetnextfilename{flowcurves}%
\input{fig_presentation/flowcurves.pgf}
\end{columns}
\vspace{-\baselineskip}\textit{\scriptsize Srour et al. Macromol. Rapid Comm., 36, 55 (2015)}
\end{frame}

\begin{frame}{Linear rheology}
	\begin{columns}
	\column{0.5\textwidth}
	\tikzsetnextfilename{linear_rheology_PImBr}%
	\input{fig_presentation/linear_rheology_PImBr.pgf}
	\[ G = \frac{c}{N}k_\mathrm{B}T \]
	\begin{description}[$N$]
	\item[$c$] monomer concentration
	\item[$N$] \# monomers between CL
	\end{description}
	\column{0.45\textwidth}
	\pause
	\begin{itemize}
	\item $N \approx 800 N_0$ !
	\item 800 chains between CL
	\end{itemize}
	\begin{flushright}\includegraphics[height=14\baselineskip]{presentation/Thaumetopea_pityocampa_01.jpg}\\
	\vspace{-1.5em}\colorbox{lightgray}{\scriptsize Wikimedia commons}
	\end{flushright}
	\end{columns}
\end{frame}

\begin{frame}{Tweak repeated ions + counterions}
	\vspace{1.5\baselineskip}
	\tikzsetnextfilename{postfunc}%
	\input{fig_presentation/postfunc.pgf}

	\bigskip
	\tikzsetnextfilename{postfunc_chem}%
	\input{fig_presentation/postfunc_chem.pgf}
\end{frame}

\begin{frame}\frametitle{Counterion condensation}
\begin{itemize}
\item Direct Hoffmeister series $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$
\item \ce{\pi+-\pi+} interactions {\tiny\setatomsep{1.5em}$\raisebox{-0.5em}{\chemfig{[2]*5(--N(-)---)}}>\raisebox{-0.5em}{\chemfig{[2]*5(-N(-)-=N-=)}}$}
\end{itemize}
\begin{tikzpicture}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\node at (0.33\columnwidth, -8\baselineskip) {PBr} [->, level distance=5em] 
	child [grow=east] {node(PImBr) {\ce{PIm+Br-}} [level distance=5em]
		child[grow=-45] {node(PImF) {\ce{PIm+F-}}} 
		child {node(PImCl) {\ce{PIm+Cl-}}} 
		child[grow=45] {node(PImI) {\ce{PIm+I-}}}} 
	child [grow=west] {node(PPyrBr) {\ce{PPyr+Br-}} [level distance=5em]
		child[grow=-135] {node(PPyrF) {\ce{PPyr+F-}}} 
		child {node(PPyrCl) {\ce{PPyr+Cl-}}} 
		child[grow=135] {node(PPyrI) {\ce{PPyr+I-}}}};
\begin{pgfonlayer}{background}
	\draw[ilmorange, line width=0.3em, ->] (PImI) to[bend right] (PImBr.north) ..controls (PImBr.center) and (PImCl.center) .. (PImCl.base) to[bend left] (PImF);
	\draw[ilmorange, line width=0.3em, ->] (PPyrI) to[bend left] (PPyrBr.north) ..controls (PPyrBr.center) and (PPyrCl.center) .. (PPyrCl.base) to[bend right] (PPyrF);
\end{pgfonlayer}
	\foreach \x in {PImF,PImCl,PPyrF}{
	\draw[ilmcolor] (\x.north west) -- (\x.base east) (\x.north east) -- (\x.base west) ;
	}
\end{tikzpicture}

\end{document}