### 2016-07-25 PALSE/talk_Kyoto.tex
funny pics for intro on interdisciplinary

\documentclass[xcolor=table]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage[super]{nth}
%\usetheme{Boadilla}
%\usecolortheme{rose}
%\usecolortheme{crane}
%\usefonttheme{structuresmallcapsserif}
%\setbeamertemplate{navigation symbols}{}

\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
\definecolor{Accent1}{rgb}{0.76,0.36,0.13}
\definecolor{Accent2}{rgb}{0.54,0.1,0.4}

\mode<presentation>{\usetheme{ilm}}
%\usecolortheme{rose}
%\useinnertheme[shadow]{circles}
%\usecolortheme{whale}
%\useoutertheme{infolines}

%\usecolortheme[named=Accent1]{structure}

%\setbeamerfont{page number in head/foot}{size=\large}
%\setbeamercolor{page number in head/foot}{fg=Main}
%% page/total
%%\setbeamertemplate{footline}[frame number]
%% pas de total
%\setbeamertemplate{footline}{%
%    	\hfill%
%	\usebeamercolor[fg]{page number in head/foot}%
%	\usebeamerfont{page number in head/foot}%
%	\insertframenumber\kern1em\vskip2pt%
%}
%
%\setbeamersize{text margin left=1em}
%\setbeamersize{text margin right=1em}

\usepackage[overlay,absolute]{textpos}
\setlength{\TPHorizModule}{10mm}
\setlength{\TPVertModule}{\TPHorizModule}
\textblockorigin{10mm}{10mm} % start everything near the top-left corner
\setlength{\parindent}{0pt}

%font
\usepackage[T1]{fontenc}
\usepackage{times}
%\usepackage[oldstylenums]{kpfonts}

%\include{macros}
\usepackage{ifthen}

%proper math and math symbols
\usepackage{amsmath,mathtools}
\usepackage{amssymb}

\usepackage{datenumber,fp}

\usepackage[separate-uncertainty = true]{siunitx}

\usepackage{tabu}
\usepackage{multirow}
\usepackage{booktabs}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, decorations.text, shapes.geometric, matrix,patterns,mindmap,fit}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.9}
\usepgfplotslibrary{fillbetween}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}
\makeatletter
\newcommand*{\overlaynumber}{\number\beamer@slideinframe}
\tikzset{
  beamer externalizing/.style={%
    execute at end picture={%
      \tikzifexternalizing{%
        \ifbeamer@anotherslide
        \pgfexternalstorecommand{\string\global\string\beamer@anotherslidetrue}%
        \fi
      }{}%
    }%
  },
  external/optimize=false
}
\let\orig@tikzsetnextfilename=\tikzsetnextfilename
\renewcommand\tikzsetnextfilename[1]{\orig@tikzsetnextfilename{#1-\overlaynumber}}
\makeatother

\tikzset{every picture/.style={beamer externalizing}}
\tikzexternalize
\tikzsetexternalprefix{fig_presentation/}
%\tikzset{external/optimize=false}
%\tikzset{external/force remake}

%link or play movies
\usepackage{multimedia}

%chemistry
\usepackage[version=3]{mhchem}
\usepackage{chemfig}
\usepackage{setspace}

%beamer related package

\usepackage{todonotes}
\presetkeys{todonotes}{inline}{}

\tikzset{onslide/.code args={<#1>#2}{%
  \only<#1>{\pgfkeysalso{#2}} %
}}%

%bibliography
\usepackage[style=authoryear-comp, language=british,eprint=false, url=false, doi=false, sortcites=true, sorting=none, isbn=false, firstinits=true,maxcitenames=6]{biblatex}
%minimal citations
\AtEveryCitekey{%
	\clearfield{title}
	\clearfield{pages}
	\clearfield{volume}
	\clearfield{number}
	\clearfield{month}}
\newcommand{\myfullcite}[1]{{\scriptsize\fullcite{#1}}}
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}
%\bibliography{biblio}

\newcolumntype{P}[1]{>{\raggedright}p{#1}}

\institute[iLM]{Univ Lyon, Université Claude Bernard Lyon 1, CNRS, Institut Lumière Matière}
\title[processionary gel]{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}
\author[M. Leocmach]{Mathieu Leocmach \hfill{\usebeamerfont{normal text}\texttt{\usebeamercolor[fg]{normal text}\footnotesize @LamSonLeoc\includegraphics[height=1em]{Twitter_Bird}}}\vspace{-\baselineskip}}
\date{%10 August 2016
}
\titlegraphic{\vspace{-\baselineskip}
	\begin{tabu}{X[c]X[c]X[c]X[c]}
		\multicolumn{2}{c}{\usebeamerfont{institute}\usebeamercolor[fg]{institute}ENS de Lyon, Laboratoire de Chimie}&
		\multicolumn{2}{c}{\usebeamerfont{institute}\usebeamercolor[fg]{institute}ENS de Lyon, Laboratoire de Physique}\\\cmidrule(r){1-2}\cmidrule(l){3-4}
		\includegraphics[height=3\baselineskip]{presentation/Hassan}&
		\includegraphics[height=3\baselineskip]{presentation/Cyrille}&
		\includegraphics[height=3\baselineskip]{presentation/Nicolas_Taberlet}&
		\includegraphics[height=3\baselineskip]{../Yaourt/Seb}\\
		Hassan & Cyrille & Nicolas & S\'{e}bastien\\
		Srour & Monnereau & Taberlet & Manneville\\\cmidrule(r){1-2}
		\multicolumn{2}{c}{Martien Duvall Deffo Ayagou}&\\
		\multicolumn{2}{c}{Thi Thanh-Tam Nguyen}&\raisebox{-0.1em}[0pt][0pt]{\includegraphics[height=2\baselineskip]{presentation/invest-avenir.jpg}}&\raisebox{-0.1em}[0pt][0pt]{\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{../Yaourt/NEW-Logo-ERC-OUTLINE}}\\
	\end{tabu}

\bigskip
Martien Duvall Deffo Ayagou, Thi Thanh-Tam Nguyen%, Chantal Andraud

	%\vfill
	%\includegraphics[height=2\baselineskip]{logo_ens-lyon}\quad
	%\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
	%\includegraphics[height=2\baselineskip]{../../Yaourt/CNRSfilaire-Q}\quad
	%\includegraphics[height=2\baselineskip]{CRPP}\quad
	%\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}
	}

\newlength{\mylength}

%\includeonly{creep_beamer}

\begin{document}
\tikzset{every mark/.append style={scale=0.8}}
\pgfplotsset{every axis/.append style={footnotesize}}

\pgfplotscreateplotcyclelist{earthy}{%
{red!40!black,mark=o},
{red!60!black,mark=triangle, every mark/.append style={rotate=180}},
{red!80!black,mark=square},
{red,mark=triangle},
{red!80!yellow, mark=diamond},
red!60!yellow,
red!40!yellow,
}

\AtBeginSection[]{
	\addtocounter{framenumber}{-1}
	\begin{frame}[plain]
		\tableofcontents[currentsection, hideothersubsections]
	\end{frame}
}

\begin{frame}{\pgfuseimage{cnrs-logo}\hspace*{0.3cm}\pgfuseimage{ucbl-logo}\pgfuseimage{univlyon-logo}}%[plain]
	\titlepage
\end{frame}

\setcounter{framenumber}{0}

\begin{frame}{What chemists/physicists think I do}
	\begin{columns}[t]
	\column{0.5\textwidth}
	\begin{flushleft}\structure{chemistry}

	``This counterion is more chaotropic''

	\vspace{\baselineskip}
	\includegraphics[width=\textwidth]{presentation/sorciere}
	\end{flushleft}
	\column{0.5\textwidth}
	\begin{flushright}\structure{rheology}

	``We impose a constant shear stress''
	\end{flushright}
	\hspace{0.25\textwidth}\includegraphics[height=10\baselineskip]{presentation/rascar}
	\end{columns}
\end{frame}

\begin{frame}{Physical chemistry}
\tikzsetnextfilename{chem_rheol}%
	\input{fig_presentation/chem_rheol.pgf}
\end{frame}

\begin{frame}{Fine chemistry}
	\vspace{\baselineskip}
	\tikzsetnextfilename{polymerisation}%
	\input{fig_presentation/polymerisation.pgf}

	\bigskip
	\tikzsetnextfilename{polymerisation_PImBr}%
	\input{fig_presentation/polymerisation_PImBr.pgf}
\end{frame}

\begin{frame}{Physical polyelectrolytes hydrogel}
	\tikzsetnextfilename{polyelectrolytes}%
\input{fig_presentation/polyelectrolytes.pgf}
\begin{columns}
\column{0.5\textwidth}
\structure{Reversible gel}
\begin{itemize}
	\item in deionised, neutral water
	\item yield stress fluid
\end{itemize}
\structure{Solution} if
\begin{itemize}
	\item ionic strength, pH
	\item non ionic head
\end{itemize}
$\Rightarrow$ Head-to-body ionic bonds

\column{0.5\textwidth}
\hfill Herschel-Bulkley law\\
\hfill $\sigma =\sigma_c + A \dot{\gamma}^n$
\tikzsetnextfilename{flowcurves}%
\input{fig_presentation/flowcurves.pgf}
\end{columns}
\vspace{-\baselineskip}\textit{\scriptsize Srour et al. Macromol. Rapid Comm., 36, 55 (2015)}
\end{frame}

\begin{frame}{Linear rheology}
	\begin{columns}
	\column{0.5\textwidth}
	\tikzsetnextfilename{linear_rheology_PImBr}%
	\input{fig_presentation/linear_rheology_PImBr.pgf}
	\[ G = \frac{c}{N}k_\mathrm{B}T \]
	\begin{description}[$N$]
	\item[$c$] monomer concentration
	\item[$N$] \# monomers between CL
	\end{description}
	\column{0.45\textwidth}
	\pause
	\begin{itemize}
	\item $N \approx 800 N_0$ !
	\item 800 chains between CL
	\end{itemize}
	\begin{flushright}\includegraphics[height=14\baselineskip]{presentation/Thaumetopea_pityocampa_01.jpg}\\
	\vspace{-1.5em}\colorbox{lightgray}{\scriptsize Wikimedia commons}
	\end{flushright}
	\end{columns}
\end{frame}

\begin{frame}{Tweaking repeated ions + counterions}
	\vspace{1.5\baselineskip}
	\tikzsetnextfilename{postfunc}%
	\input{fig_presentation/postfunc.pgf}

	\bigskip
	\tikzsetnextfilename{postfunc_chem}%
	\input{fig_presentation/postfunc_chem.pgf}
\end{frame}

\begin{frame}<1>[label=condensationranking]{Counterion condensation}
	\tikzsetnextfilename{solubility}%
	\input{fig_presentation/solubility.pgf}
	\begin{itemize}
		\item Too much condensation prevents dissolution
		\begin{itemize}
			\item water poor solvent for polymer backbone
			\item polyelectrolytes solubilized by electrostatic repulsion
		\end{itemize}
		\item<2-> Direct Hoffmeister series $\ce{I-}\alert{<}\ce{Br-}\alert{<}\ce{Cl-}\alert{<}\ce{F-}$\\
	\hfill\textit{\scriptsize Hofmeister 1888; Zhang Annu Rev Phys Chem 2010; Schwierz Langmuir 2010}
		\item<3-> \ce{\pi+-\pi+} interactions trap counterions\raisebox{-0.5em}{%
			\begin{tikzpicture}[every node/.style={inner ysep=0,font=\tiny\setatomsep{1.5em}}]
			\node[left=0.5em] (Pyr){\chemfig{[0]*5(--N(-)---)}};
			\node[right=0.5em] (Im){\chemfig{[2]*5(-N(-)-=N-=)}};
			\node[font=\normalsize]{$\alert{<}$};
			\end{tikzpicture}}\\
			\hfill\textit{\scriptsize
			Geronimo PCCP 2011;
			Matthews PCCP 2014;
			Gao J Mol Graph Model 2015}
	\end{itemize}
\end{frame}

\begin{frame}{Condensation and solubility}
	\begin{itemize}
		\item All counterions condensed $\Rightarrow$ neutral $\Rightarrow$ not soluble
	\end{itemize}
	\tikzsetnextfilename{polyelectrolytes_condensed}%
	\input{fig_presentation/polyelectrolytes_condensed.pgf}
	\begin{itemize}
		\item ``Enough'' free counterions $\Rightarrow$ repulsion $\Rightarrow$ soluble
	\end{itemize}
	\tikzsetnextfilename{polyelectrolytes}%
	\input{fig_presentation/polyelectrolytes.pgf}
\end{frame}

\againframe<2-3>{condensationranking}

\begin{frame}{\# chains between cross links}%
	\tikzsetnextfilename{frequency}%
	\input{fig_presentation/frequency.pgf}
	\begin{itemize}
	\item Procession length increases with counterion condensation
	\end{itemize}
\end{frame}

\begin{frame}{Polyelectrolytes physics}
	\vspace{\baselineskip}
	\structure{Can we measure counterion condensation from rheology?}

	\begin{itemize}
		\item 1 dissociated counterion every \alert{$A$} monomers
		\begin{itemize}\item large $A$, more condensed\end{itemize}
		\item poor solvent: reduced temperature $\alert{\tau} = 1 -T/\Theta$
		\begin{itemize}\item \ce{PIm+F-} non soluble at \SI{100}{\celsius} $\Rightarrow\tau>0.2$ \end{itemize}
	\end{itemize}
	\alert{Scaling theory}\hfill \textit{\scriptsize Rubinstein Macromol. 1996}
	\tikzsetnextfilename{scales}%
	\input{fig_presentation/scales.pgf}

	\begin{itemize}
		\item extension parameter $\alert{B} = \frac{\text{extended length}}{\text{diameter}}$ of electrostatic blob
		\smallskip
		\item scales $\geq D$ depend only on $B \sim A^{4/3} \tau$
	\end{itemize}

	\begin{center}
	\structure{Can we measure $B$ from rheology?}
	\end{center}
\end{frame}

\begin{frame}<1>[label=endlinear]{Stretching a polyelectrolyte chain}
	\vspace{\baselineskip}
	\begin{itemize}
		\item Unfold largest scales first\hfill \textit{\scriptsize Pincus Macromol. 1976}
	\end{itemize}

	\tikzsetnextfilename{scales_stretch}%
	\begin{tikzpicture}
	\node[inner sep=0](scales){\input{fig_presentation/scales.pgf}};
	\draw[->, ultra thick, ilmcolor] (scales.south east) ++(0,-0.3em) -- +(-0.33\textwidth,0) node[midway, below] {linear};
	\draw[->, ultra thick,ilmorange] (scales.south) ++(0,-0.3em) -- +(-0.35\textwidth,0) node[midway, below] {non linear};
	\end{tikzpicture}

	\pause
	\begin{itemize}
		\item narrow linear regime $\Rightarrow R_0\approx r_\mathrm{scr} \Rightarrow \left(\frac{B}{B_0}\right)^3 = 1 + \frac{B}{B_s}$
		\begin{itemize}
			\item \alert{$B_s$} screening from \ce{-PO_3^{2-}} heads\tikzsetnextfilename{head}
				\begin{tikzpicture}
					\let\mrad\relax%
					\newlength\mrad%
					\setlength{\mrad}{1ex}%
					%face
					\input{fig_presentation/head.pgf}
					%% horns
					\draw[every node/.style={shape=circle,inner sep=0.2\mrad,shading=radial,inner color=white!50!ilmcolor,outer color= ilmcolor!90!black}] (emoticon.70) -- ++(70:0.4\mrad) node{} (emoticon.110) -- ++(110:0.4\mrad) node{};
				\end{tikzpicture}
			\item if $B \ll B_s$ screening from uncondensed counterions \[B = B_0 \sim bc \left(\frac{k_\mathrm{B}T}{G}\right)^{2/3}\]
			\item if $B \gg B_s$ screening from heads: $B = \left(B_0^3/B_s\right)^{1/2} > B_0$
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}<1>[label=strainsweep]{Strain sweep}
	\vspace{0.5\baselineskip}
	\tikzsetnextfilename{strain}%
	\input{fig_presentation/strain.pgf}

	\vspace{-0.5\baselineskip}
	\begin{overprint}%
	\onslide<-2>%
	\begin{itemize}
		\item narrow linear regime $\gamma_0\ll 1$
		\item<2> extension parameter increases with condensation
	\end{itemize}
	\onslide<3->%
	\begin{itemize}
		\item bonding energy increases with $B$
		\item<4> measure of counterion condensation $A$
	\end{itemize}
	\end{overprint}
\end{frame}

\againframe<2>{endlinear}
\againframe<2>{strainsweep}

\begin{frame}{Extension of the electrostatic blobs}
\vspace{\baselineskip}

\tikzsetnextfilename{stretch}%
	\tikzsetnextfilename{stretch}%
	\input{fig_presentation/stretch.pgf}
	\[E_\mathrm{stretch} = \text{area increment}(\gamma,B) \times \text{surface tension}(\tau,B)\]

	\begin{itemize}
		\item fluidisation = rupture $\Rightarrow E_\mathrm{stretch}(\gamma_c) = E_\text{head-body}$
		\begin{itemize}
			\item \structure{$\tau(E_\text{head-body}, B, \gamma_c)$}
		\end{itemize}
		\item ionic bond energy depends on local $\epsilon_r$%\\\hfill\textit{\scriptsize Rehm Chem Soc Rev. 2010}
		\begin{itemize}
			\item in water $E_\text{head-body}\approx 2k_\mathrm{B}T$
			\item large $B\Rightarrow$ low $\epsilon_r$ $\Rightarrow$ high $E_\text{head-body}$
		\end{itemize}
		\item smallest $B$ is \ce{PPyr+I-} $\Rightarrow \textcolor{ilmcolor}{\tau = 0.4\pm0.07}, \Theta\approx\SI{220}{\celsius}$
	\end{itemize}
\end{frame}

\againframe<3-4>{strainsweep}

\begin{frame}{Conclusion}
	\begin{itemize}
		\item measure of microscopic parameters from rheology
		\begin{itemize}
			\item counterion condensation
			\item head-to-body bonding energy
		\end{itemize}
		\item post-functionalisation allows to change $A\times 100$
		\begin{itemize}
			\item \ce{PPyr+I-}: 10 free counterions per polymer
			\item \ce{PIm+Br-}: 1 free counterion every 9 polymers
		\end{itemize}
		\item mechanical properties $\times 1000$
		\begin{itemize}
			\item \ce{PPyr+I-}: $G\approx\SI{10}{\kilo\pascal}$, fluidizes at $\gamma_c\approx 2\%$
			\item \ce{PIm+Br-}: $G\approx\SI{10}{\pascal}$, fluidizes at $\gamma_c\approx 800\%$
		\end{itemize}
		\item highly tunable hydrogels from short, linear polymer chains
		\begin{itemize}
			\item no additive
			\item yield stress fluids
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Processionary phase diagram}
	\begin{columns}
	\column{0.65\textwidth}
	\vspace{\baselineskip}
	\tikzsetnextfilename{phase_diag}%
	\input{fig_presentation/phase_diag.pgf}
	\column{0.35\textwidth}
	\begin{flushright}\includegraphics[height=14\baselineskip]{presentation/Thaumetopea_pityocampa_01.jpg}\\
	\vspace{-1.5em}\colorbox{lightgray}{\scriptsize Wikimedia commons}
	\end{flushright}
	\end{columns}
\end{frame}

\end{document}